---
title: "Final Project Analysis"
author: "Charles Yang, Isaac Tegeler, Carter Wunsch, Bennet Brosseau, Nicholas Heger"
date: "10/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

# Data was stored in CSV (Comma Delimited), but was tab delimited.
data <- read.table("VIQT_data.csv", sep="\t", header=TRUE)
```

# Exploratory Analysis
### Charles S1-6, education v. Full Score
```{r}
data %>%
  group_by(S1) %>%
  summarise(`Full Score`=mean(score_full), .groups="drop") %>%
  filter(S1 > 0) %>%
  ggplot(aes(x=S1, y=`Full Score`)) +
  geom_bar(stat="identity", width=.7) +
  ggtitle("Mean Full Scores by S1 Response")
```

```{r}
data %>%
  group_by(S2) %>%
  summarise(`Full Score`=mean(score_full), .groups="drop") %>%
  filter(S2 > 0) %>%
  ggplot(aes(x=S2, y=`Full Score`)) +
  geom_bar(stat="identity", width=.7) +
  ggtitle("Mean Full Scores by S2 Response")
```

```{r}
data %>%
  group_by(S3) %>%
  summarise(`Full Score`=mean(score_full), .groups="drop") %>%
  filter(S3 > 0) %>%
  ggplot(aes(x=S3, y=`Full Score`)) +
  geom_bar(stat="identity", width=.7) +
  ggtitle("Mean Full Scores by S3 Response")
```

```{r}
data %>%
  group_by(S4) %>%
  summarise(`Full Score`=mean(score_full), .groups="drop") %>%
  filter(S4 > 0) %>%
  ggplot(aes(x=S4, y=`Full Score`)) +
  geom_bar(stat="identity", width=.7) +
  ggtitle("Mean Full Scores by S4 Response")
```

```{r}
data %>%
  group_by(S5) %>%
  summarise(`Full Score`=mean(score_full), .groups="drop") %>%
  filter(S5 > 0) %>%
  ggplot(aes(x=S5, y=`Full Score`)) +
  geom_bar(stat="identity", width=.7) +
  ggtitle("Mean Full Scores by S5 Response")

```

```{r}
data %>%
  group_by(S6) %>%
  summarise(`Full Score`=mean(score_full), .groups="drop") %>%
  filter(S6 > 0) %>%
  ggplot(aes(x=S6, y=`Full Score`)) +
  geom_bar(stat="identity", width=.7) +
  ggtitle("Mean Full Scores by S6 Response")
```

```{r}
data$education <- recode(data$education, `1`="< HS", `2`="HS", `3`="UNI", `4`="GRAD") %>% factor(levels=c("< HS", "HS", "UNI", "GRAD"))
data %>%
  group_by(education) %>%
  summarise(`Full Score`=mean(score_full), .groups="drop") %>%
  ggplot(aes(x=education, y=`Full Score`)) +
  geom_bar(stat="identity", width=.7) +
  ggtitle("Mean Full Scores by Education")
```

### Isaac S7 - 12, urban v. Full Score

We will first look a the impact of specific questions on the final score to see if there are any questons that were particullarly harder than the others.
```{r}
#Answer distribution for questions 7 - 12
x <- data %>% select(Q7, Q8, Q9, Q10, Q11, Q12) %>%
  pivot_longer(cols=Q7:Q12, names_to="question", values_to="answer")
x$question = factor(x$question, levels=c("Q7", "Q8", "Q9", "Q10", "Q11", "Q12"))
ggplot(x, aes(x=answer)) + geom_histogram(binwidth = 1) + facet_wrap(~question) + ggtitle("Which questons are the hardest?")
```

Question 7 appears to be the hardest to answer as it has the largest distribution of responses. The question asked which two words have the same meaning from the following list: [ deal claim plea recoup sale ]. We also see that people are much more likely to answer this questions rather than say they don't know (there are fewer answers of 0 which indicates they did not answer the question).

Next we will see if how urban of a location you live impacts the final score.

```{r}
data$urban <- recode(data$urban, `1`="Rural", `2`="Suburban", `3`="Urban", ) %>% factor(levels=c("Rural", "Suburban", "Urban"))

ggplot(data, aes(x=score_full)) +geom_histogram() +facet_wrap(~urban, scales="free_y")
```

The distributions show that students in Rural and Suburban areas have similar levels of vocabulary while those in urban areas have weaker vocuabularies. Where there are even some in the rural areas that scored negativly on the test while there are none who did so in suburban and rural areas.

### Nicholas S13 - 18, gender v. Full Score
```{r}

```

### Carter S19 - 24, engnat v. Full Score
```{r}

```

### Bennet S25-30, age v. Full Score
### S25 - I wish people were more spiritual.
```{r}
data %>%
  group_by(S25) %>%
  summarise(`Full Score`=mean(score_full), .groups="drop") %>%
  filter(S25 > 0) %>%
  ggplot(aes(x=S25, y=`Full Score`)) +
  geom_bar(stat="identity", width=.7) +
  ggtitle("Mean Full Scores by S25 Response")
```


### S26 - I always do the bare minimum I need to get by.
```{r}
data %>%
  group_by(S26) %>%
  summarise(`Full Score`=mean(score_full), .groups="drop") %>%
  filter(S26 > 0) %>%
  ggplot(aes(x=S26, y=`Full Score`)) +
  geom_bar(stat="identity", width=.7) +
  ggtitle("Mean Full Scores by S26 Response")
```


### S27 - I like to play devils advocate.
```{r}
data %>%
  group_by(S27) %>%
  summarise(`Full Score`=mean(score_full), .groups="drop") %>%
  filter(S27 > 0) %>%
  ggplot(aes(x=S27, y=`Full Score`)) +
  geom_bar(stat="identity", width=.7) +
  ggtitle("Mean Full Scores by S27 Response")
```


### S28 - I am not bothered by messy people.
```{r}
data %>%
  group_by(S28) %>%
  summarise(`Full Score`=mean(score_full), .groups="drop") %>%
  filter(S28 > 0) %>%
  ggplot(aes(x=S28, y=`Full Score`)) +
  geom_bar(stat="identity", width=.7) +
  ggtitle("Mean Full Scores by S28 Response")
```


### S29 - I am more artistic than scientific.
```{r}
data %>%
  group_by(S29) %>%
  summarise(`Full Score`=mean(score_full), .groups="drop") %>%
  filter(S29 > 0) %>%
  ggplot(aes(x=S29, y=`Full Score`)) +
  geom_bar(stat="identity", width=.7) +
  ggtitle("Mean Full Scores by S29 Response")

```


### S30 - I am not quite sure what I want.
```{r}
data %>%
  group_by(S30) %>%
  summarise(`Full Score`=mean(score_full), .groups="drop") %>%
  filter(S30 > 0) %>%
  ggplot(aes(x=S30, y=`Full Score`)) +
  geom_bar(stat="identity", width=.7) +
  ggtitle("Mean Full Scores by S30 Response")
```


### Mean Full Scores by Age
```{r}
data %>%
  group_by(age) %>%
  summarise(`Full Score`=mean(score_full), .groups="drop") %>%
  filter(age < 100) %>%
  ggplot(aes(x=age, y=`Full Score`)) +
  geom_bar(stat="identity", width=.7) +
  ggtitle("Mean Full Scores by Age")
```
